---
title: "multidrug"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Multi-Drug Resistance Data

source:https://discover.nci.nih.gov/nature2000/natureintromain.jsp

## Drug activity data

Copy the url to the "gz format" of drug activity data and read the data:

```{r}
url_da <- "https://discover.nci.nih.gov/nature2000/data/Archives/a_matrix1400.txt.gz"
multidrug_compound_raw <- data.table::fread(url_da, header = TRUE)
```

From description, the columns are:

**Drug activity data (A-matrix).**
Tab delimited (1400 drugs x 60 cells) Used to generate dendrogram in Figure 2b
Columns named in Excel format.
  A:	NSC number (i.e. a unique drug identifier)
  B:	Drug name, if known. Else, the NSC number.
  C:	Mechanism of action, if known.
  D-BK:	Drug activities (-Log10 Gi50)..

Therefore, we will save it as an assay with A:C as `rowData`:

```{r}
multidrug_compound_raw[,-1:-3] %>% class()
```

```{r}
library(data.table)
library(SummarizedExperiment)
## drop the 2nd and 3rd column and keep the first one as rownames
cmp <- multidrug_compound_raw[,-2:-3] %>% data.frame(row.names = 1) %>% as.matrix()
rowdata <- multidrug_compound_raw[,2:3] %>%  DataFrame()
compound <- SummarizedExperiment(assays = list(compound = cmp), rowData = rowdata,
                                 colData = list(Tissue2 = factor(substring(colnames(cmp), 0, 2))))
```



## Gene expression data

Expression of 1375 genes ( a bigger dataset is also available):

```{r}
url_ge <- "https://discover.nci.nih.gov/nature2000/data/Archives/t_matrix1375.txt.gz"
multidrug_exp_raw <- data.table::fread(url_ge, header = TRUE)
```

Description: 

**Gene expression data description (selected genes, 1375)**
Tab delimited (1375 genes x 60 cells). Columns named in Excel format.
  A:	IMAGE Clone ID
  B:	Gene description.
  C:	5ACC: 5' GenBank accession number.
  D:	3ACC: 3' GenBank accession number.
  E-BL:	Gene expression levels expressed as Log2(ratio), where ratio = the red/green fluorescence ration after computational balancing of the two channels.

We will drop A, C, D:
```{r}
expression <- multidrug_exp_raw[,-c(1,3:4)] %>% data.frame(row.names = 1) %>% as.matrix()
expression <- SummarizedExperiment(assays = list(expression = expression ), 
                                   colData = list(Tissue = factor(substring(colnames(expression), 0, 2))))
```

## ABC genes

ABC transport genes from https://discover.nci.nih.gov/abc/2004_cancercell_abstract.jsp#supplement:

Copy the link to the txt file of Supplemental Table S1
```{r}
url_abc <- "https://discover.nci.nih.gov/abc/2004_cancercell_TableS1.txt"
multidrug_abc_raw <- data.table::fread(url_abc, header = TRUE)
multidrug_abc <- multidrug_abc_raw %>% data.frame(row.names = 1) %>% as.matrix()
```

The rownames are gene names and colnames include sample names and phenotypes, but do not match mixOmics' multidrug sample names.
Leaving it at this point for now.

The column names are not necessarily in the same order so I'm wondering if every assay has its
## MultiAssayExperiment

```{r}
testthat::expect_identical(colnames(expression), colnames(compound))
colnames(expression) %in% colnames(compound)
colnames(compound) %in% colnames(expression)
sapply(identical(colnames(expression), colnames(multidrug_abc)), FUN = identical, colnames(compound))
```

```{r}
exps <- ExperimentList(list(Expression = expression, Compound = compound, ABC.trans = multidrug_abc))
multimae <- MultiAssayExperiment(experiments = exps)
```

The colData from SE assays are not automatically included in MAE.

For now I'll proceed with the tidy liver toxicity for pkg enhancement because
I really need to get something done!

What we have to do:

  * see how order of samples in individual assays affect colData order?
  * remove redundanct colData in liver.toxicity (Treatment.Group and Dose.Group)
  * Get a better sense of MultiAssayExperiment ecosystem and how it can be leveraged
  using its getters
  * Create a better dataset with colData and rowData and multiple omics]
  * Possibly also include a SingleCell MAE
  * Make data creation procedure fully transparent and reproducible.
